---
title: "Naive Bayes Assignment 4"
author: "Ridhwan Firdaus bin Mohd Jata"
date: "3/16/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the library
Listed below is the library that we will be using
```{r}
library(e1071)
library(mlr)
```

##The Dataset
We will be using the Titanic dataset
```{r}
data("Titanic")
df=as.data.frame(Titanic)
head(df)
repeating_sequence=rep.int(seq_len(nrow(df)), df$Freq)
titanic=df[repeating_sequence,]
titanic$Freq = NULL
head(titanic)
```


## Fit the dataset into our Naive Bayes model
```{r}
naivebayesmodel=naiveBayes(Survived ~., data=titanic)
naivebayesmodel
```

## Prediction on the dataset 
```{r}
predict=predict(naivebayesmodel,titanic)
```

## Confusion matrix to check accuracy
```{r}
table(predict,titanic$Survived)
```

##Create a classification task on Titanic dataset and specify the target feature
```{r}
task = makeClassifTask(data = titanic, target = "Survived")
```

##Initialize the Naive Bayes classifier
```{r}
selectedmodel = makeLearner("classif.naiveBayes")
```

##Train the model
```{r}
NB_mlr = train(selectedmodel, task)
```

## Read the model learn results
```{r}
NB_mlr$learner.model
```

#Predict on the dataset without passing the target feature
```{r}
predictionsmlr = as.data.frame(predict(NB_mlr, newdata = titanic[,1:3]))
```

##Confusion matrix to check accuracy 
```{r}
confusionmatrix <- table(predictionsmlr[,1],titanic$Survived)
confusionmatrix
```

## Accuracy ,sensitivity and specificity

```{r}
accuracy <- round(sum(diag(confusionmatrix))/sum(confusionmatrix),2)
sprintf('Accuracy is %s', accuracy)
sensitivity <- round(349/(126 + 349), 2)
sprintf('Sensitivity is %s', sensitivity)
specificity <- round(1364/(1364 + 362),2)
sprintf('Specificity is %s', specificity)
```

