---
title: "Assignment 5 eigenvalue and eigenvector decomposition"
author: "Ridhwan Firdaus bin Mohd Jata"
date: "3/30/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(magrittr) # needs to be run every time you start R and want to use %>%
library(dplyr)
library(psych)
```

## Data used for eigenvalue and eigenvector decomposition
In this section the data used for eigenvalue and eigenvector decomposition will be from kaggle. The link is attached here https://www.kaggle.com/abcsds/pokemon . Below is the structure of the data

```{r}
data <- read.csv('Pokemon.csv')
data_pca <- data %>% select(HP, Attack, Defense, Speed, Sp..Atk, Sp..Def)
summary(data_pca)
```
```{r}
head(data_pca)
```

## Placing the data on a square matrix
The raw data is imported and then a correlation matrix is generated (this is using simulated data based on the original correlation matrix).

```{r}
matrix<-cor(data_pca)
matrix
```
The next step is to find out wether the data can be reduced or not. One basic test is Bartlett’s test of sphericity – the null hypothesis of the test is that the correlation matrix is an identity matrix – or that the matrix has one’s on the diagonal and zeroes on all the off diagonals.
```{r}
cortest.bartlett(matrix, n = nrow(data))
```
Test’s results agree with the correlation matrix form, it is straightforward that variables are positively correlated with each other.Next step is to solve the eigenvalues and eigenvectors from the correlation matrix
```{r}
eigen<-eigen(matrix) 
L<-eigen$values
Vm<-matrix(0,nrow=ncol(data_pca),ncol=ncol(data_pca))
diag(Vm)<-L
Vm
```
Therefore listed above is the eigenvalue of the correlation matrix

