---
title: "Assignment 7 Graph Theory"
author: "Ridhwan Firdaus bin Mohd Jata"
date: "3/31/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
The two primary aspects of networks are a multitude of separate entities and the connections between them.The entities are referred to as nodes or vertices of a graph, while the connections are edges or links

## Data used
The data sets that will be used relates to French trains; it contains aggregate daily total trips per connecting stations.
```{r}
library(readr)

url <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-02-26/small_trains.csv"
small_trains <- read_csv(url)
head(small_trains)
```

## Data Cleaning and preparation 
Data will be changed in order to create a new category called routes. It contains a single entry for each connected station. It also includes the average journey time it takes to go between stations.
```{r}
library(dplyr)

routes <- small_trains %>%  group_by(departure_station, arrival_station) %>%  summarise(journey_time = mean(journey_time_avg)) %>%  ungroup()%>%  mutate(from = departure_station,to = arrival_station) %>%  select(from, to, journey_time)

head(routes)

```
## Creating the adjency matrix
```{r}
adjmatrix <- subset(routes, select = -c(journey_time))
head(adjmatrix)
edges_mat = matrix(nrow = length(adjmatrix$from), ncol = length(adjmatrix$to), data=0, dimnames = list(adjmatrix$from, adjmatrix$to))
edges_mat[as.matrix(adjmatrix)] = 1
head(edges_mat)
```

## Transform data into graph table
In order to prepare routes for this transformation, it has to contain two variables specifically named from and to.The  member of the group is called a node (or vertex) in the graph, and a relationship between nodes is called an edge.
```{r}
library(tidygraph)

graph_routes <- as_tbl_graph(routes)

graph_routes
library(stringr)

graph_routes <- graph_routes %>%  activate(nodes) %>%  mutate(    title = str_to_title(name),    label = str_replace_all(title, " ", "\n")    )
stations <- graph_routes %>%  activate(nodes) %>%  pull(title)

```
## Visualizing
In this section we will visualize all the nodes and edges from the dataset
```{r}
library(ggplot2)
library(ggraph)
thm <- theme_minimal() +
  theme(
    legend.position = "none",
     axis.title = element_blank(),
     axis.text = element_blank(),
     panel.grid = element_blank(),
     panel.grid.major = element_blank(),
  ) 

theme_set(thm)
graph_routes %>%  ggraph(layout = "kk") + geom_node_text(aes(label = label, color = name), size = 3) +geom_edge_diagonal(color = "gray", alpha = 0.4) 
```
## Finding the shortest distance
In this section we will identify the shortest route from station Arras to station Nancy 
```{r}
from <- which(stations == "Arras")
to <-  which(stations == "Nancy")

shortest <- graph_routes %>%  morph(to_shortest_path, from, to, weights = journey_time)
shortest %>%  mutate(selected_node = TRUE) %>%  unmorph()
shortest <- shortest %>%   mutate(selected_node = TRUE) %>%   activate(edges) %>%   mutate(selected_edge = TRUE) %>%   unmorph() 
shortest <- shortest %>%  activate(nodes) %>%  mutate(selected_node = ifelse(is.na(selected_node), 1, 2)) %>%  activate(edges) %>%  mutate(selected_edge = ifelse(is.na(selected_edge), 1, 2)) %>%  arrange(selected_edge)
```
## Visualizing the shortest distance
```{r}
shortest %>%
  ggraph(layout = "kk") +
    geom_edge_diagonal(aes(alpha = selected_edge), color = "gray") +
    geom_node_text(aes(label = label, color =name, alpha = selected_node ), size = 3) 
```
## Summarizing the graph table
```{r}
shortest %>%
  activate(edges) %>%
  filter(selected_edge == 2) %>%
  as_tibble() %>%
  summarise(
    total_stops = n() - 1,
    total_time = round(sum(journey_time) / 60)
    )
```

As we can see there will be a total of 8 stops between Arras and Nancy and the total time is 23 minutes