---
title: "Logistic Regression and Correlation Matrix"
author: "Ridhwan Firdaus bin Mohd Jata"
date: "3/11/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(caTools)
```

##Dataset 
In this investigation, we will be using the diabetes dataset. Below is the structure of the dataset
```{r}
df = read.csv('diabetes.csv', header = TRUE)
str(df)
head(df)
```

## Creating Age category column 
We will classify all ages into certain category column
```{r}
df$AgeCat <- ifelse(df$Age < 21, "<21", 
                   ifelse((df$Age>=21) & (df$Age<=25), "21-25", 
                   ifelse((df$Age>25) & (df$Age<=30), "25-30",
                   ifelse((df$Age>30) & (df$Age<=35), "30-35",
                   ifelse((df$Age>35) & (df$Age<=40), "35-40",
                   ifelse((df$Age>40) & (df$Age<=50), "40-50",
                   ifelse((df$Age>50) & (df$Age<=60), "50-60",">60")))))))
df$AgeCat <- factor(df$AgeCat, levels = c('<21','21-25','25-30','30-35','35-40','40-50','50-60','>60'))
table(df$AgeCat)
head(df)
```

## Spliting dataset into train and test dataset
```{r}
require(caTools)
set.seed(3)
split = sample.split(df$Outcome,SplitRatio = 0.75)
train = subset(df, split== TRUE)
test = subset(df, split== FALSE)
head(train)
```
```{r}
nrow(df)
nrow(train)
nrow(test)
table(df$Outcome)
```

##Baseline accuracy
```{r}
baseline <- round(500/nrow(df),2)
baseline
```

Do not select a model whose accuracy is lower than 0.65

## Fit Model using all independent variables
```{r}
model = glm(Outcome~ ., data = train, family = binomial)
summary(model)
```
## Make predictions on test set
```{r}
Predict <- predict(model,type = "response" , newdata = test)
```

## Produce Confusion matrix based on the predictions
```{r}
test_tab <- table(test$Outcome, Predict > 0.5)
test_tab
accuracy_test <- round(sum(diag(test_tab))/sum(test_tab),2)
sprintf("Accuracy on test set is %s", accuracy_test)
```

 





